
```{r }
# Flight Delay

TrainData = read.csv("./input/train.csv")
TestData = read.csv("./input/test.csv")


# Data Cleaning Task

train.class = rep(0, nrow(TrainData))

for (i in 1:nrow(TrainData)){
  TrainData$ARRIVAL_DELAY[i] = ifelse(TrainData$ARRIVAL_DELAY[i] >= 10, 1, 0)
}


# test.class = rep(0, nrow(TestData))

for (i in 1:nrow(TestData)){
  TestData$ARRIVAL_DELAY[i] = ifelse(TestData$ARRIVAL_DELAY[i] >= 10, 1, 0)
}

# train.class = as.factor(as.numeric(TrainData$ARRIVAL_DELAY[i] >= 10))
# test.class = as.factor(as.numeric(TestData$ARRIVAL_DELAY[i] >= 10))


# Clean the time attributes
TrainData$SCHEDULED_ARRIVAL_HOUR = rep(0, nrow(TrainData))
TestData$SCHEDULED_ARRIVAL_HOUR = rep(0, nrow(TestData))

TrainData$SCHEDULED_DEPARTURE_HOUR =rep(0, nrow(TrainData))
TestData$SCHEDULED_DEPARTURE_HOUR = rep(0, nrow(TestData))



for (i in 1:nrow(TrainData)){
  if (TrainData$SCHEDULED_DEPARTURE[i] < 10)
  {
    TrainData$SCHEDULED_DEPARTURE_HOUR[i] = 0
  }
  else if(TrainData$SCHEDULED_DEPARTURE[i] >= 10 & TrainData$SCHEDULED_DEPARTURE[i] <100){
    TrainData$SCHEDULED_DEPARTURE_HOUR[i] = TrainData$SCHEDULED_DEPARTURE[i] %/% 10
  }
  else{
    TrainData$SCHEDULED_DEPARTURE_HOUR[i] = TrainData$SCHEDULED_DEPARTURE[i] %/% 100
  }
  
  
  
  if (TrainData$SCHEDULED_ARRIVAL[i] < 10)
  {
    TrainData$SCHEDULED_ARRIVAL_HOUR[i] = 0
  }
  else if(TrainData$SCHEDULED_ARRIVAL[i] >= 10 & TrainData$SCHEDULED_ARRIVAL[i] <100){
    TrainData$SCHEDULED_ARRIVAL_HOUR[i] = TrainData$SCHEDULED_ARRIVAL[i] %/% 10
  }
  else{
    TrainData$SCHEDULED_ARRIVAL_HOUR[i] = TrainData$SCHEDULED_ARRIVAL[i] %/% 100
  }
  
}


for (i in 1:nrow(TestData)){
  if (TestData$SCHEDULED_DEPARTURE[i] < 10)
  {
    TestData$SCHEDULED_DEPARTURE_HOUR[i] = 0
  }
  else if(TestData$SCHEDULED_DEPARTURE[i] >= 10 & TestData$SCHEDULED_DEPARTURE[i] <100){
    TestData$SCHEDULED_DEPARTURE_HOUR[i] = TestData$SCHEDULED_DEPARTURE[i] %/% 10
  }
  else{
    TestData$SCHEDULED_DEPARTURE_HOUR[i] = TestData$SCHEDULED_DEPARTURE[i] %/% 100
  }
  
  
  
  if (TestData$SCHEDULED_ARRIVAL[i] < 10)
  {
    TestData$SCHEDULED_ARRIVAL_HOUR[i] = 0
  }
  else if(TestData$SCHEDULED_ARRIVAL[i] >= 10 & TestData$SCHEDULED_ARRIVAL[i] <100){
    TestData$SCHEDULED_ARRIVAL_HOUR[i] = TestData$SCHEDULED_ARRIVAL[i] %/% 10
  }
  else{
    TestData$SCHEDULED_ARRIVAL_HOUR[i] = TestData$SCHEDULED_ARRIVAL[i] %/% 100
  }
  
}

atts.train = c("MONTH", 'DAY_OF_WEEK', "SCHEDULED_DEPARTURE", "DISTANCE",
               "SCHEDULED_DEPARTURE_HOUR", "SCHEDULED_ARRIVAL_HOUR", "ARRIVAL_DELAY")

atts.test = c("MONTH", 'DAY_OF_WEEK', "SCHEDULED_DEPARTURE", "DISTANCE","SCHEDULED_DEPARTURE_HOUR",
              "SCHEDULED_ARRIVAL_HOUR")

train.data = TrainData[atts.train]
test.data = TestData[atts.test]

train.data$ARRIVAL_DELAY = as.factor(train.data$ARRIVAL_DELAY)

t = train.data[1:30000, ]

library(party)
library(tree)
library(rpart)
library(randomForest)
# Implement Random forest
t2 = train.data[30001:45000, ]

# train.rf = randomForest(ARRIVAL_DELAY ~. ,data = train.data, importance = T, proxiimity = T)
# train.t = rpart(ARRIVAL_DELAY ~ DISTANCE + SCHEDULED_DEPARTURE_HOUR, data = train.data)
train.rf <- randomForest(ARRIVAL_DELAY ~ ., data=train.data, importance=TRUE,
                         proximity=TRUE, ntree = 300)
p = predict(train.rf, test.data)

predict = data.frame(ID = 0:(length(p) - 1), ARRIVAL_DELAY = p)
write.csv(predict, "./output/submission.csv", row.names = F)

#  train.rf <- randomForest(ARRIVAL_DELAY ~ ., data=t, importance=TRUE,
#                           proximity=TRUE, ntree = 300)
#  p = predict(train.rf, t2)
#  
#  predict = data.frame(ID = 0:(length(p) - 1), ARRIVAL_DELAY = p)



dh = summarise(group_by(t, SCHEDULED_DEPARTURE_HOUR, ARRIVAL_DELAY), freq = n())

frq_dh = rep(0,length(unique(dh$SCHEDULED_DEPARTURE_HOUR)))


for (i in unique(dh$SCHEDULED_DEPARTURE_HOUR)) {
  s = sum(dh[dh$SCHEDULED_DEPARTURE_HOUR == i, ]$freq)
  
  s_0 = dh[dh$SCHEDULED_DEPARTURE_HOUR == i & dh$ARRIVAL_DELAY == 0, ]$freq / s
  s_1 = dh[dh$SCHEDULED_DEPARTURE_HOUR == i & dh$ARRIVAL_DELAY == 1, ]$freq / s
  
  frq_dh[dh$SCHEDULED_DEPARTURE_HOUR == i & dh$ARRIVAL_DELAY == 0] = s_0
  frq_dh[dh$SCHEDULED_DEPARTURE_HOUR == i & dh$ARRIVAL_DELAY == 1] = s_1
  
}

dh['freq_dh'] = frq_dh


g1 = ggplot(dh, aes(x = SCHEDULED_DEPARTURE_HOUR, y = frq_dh, fill = ARRIVAL_DELAY)) + geom_bar( stat = "identity", alpha = 0.4)


ah = summarise(group_by(t, SCHEDULED_ARRIVAL_HOUR, ARRIVAL_DELAY), freq = n())

frq_ah = rep(0,length(unique(ah$SCHEDULED_ARRIVAL_HOUR)))


for (i in unique(ah$SCHEDULED_ARRIVAL_HOUR)) {
  s = sum(ah[ah$SCHEDULED_ARRIVAL_HOUR == i, ]$freq)
  
  s_0 = ah[ah$SCHEDULED_ARRIVAL_HOUR == i & ah$ARRIVAL_DELAY == 0, ]$freq / s
  s_1 = ah[ah$SCHEDULED_ARRIVAL_HOUR == i & ah$ARRIVAL_DELAY == 1, ]$freq / s
  
  frq_ah[ah$SCHEDULED_ARRIVAL_HOUR == i & ah$ARRIVAL_DELAY == 0] = s_0
  frq_ah[ah$SCHEDULED_ARRIVAL_HOUR == i & ah$ARRIVAL_DELAY == 1] = s_1
  
}

ah['frq_ah'] = frq_ah


g1 = ggplot(dh, aes(x = SCHEDULED_DEPARTURE_HOUR, y = frq_dh, fill = ARRIVAL_DELAY)) + geom_bar( stat = "identity", alpha = 0.4)
g2 = ggplot(ah, aes(x = SCHEDULED_ARRIVAL_HOUR, y = frq_ah, fill = ARRIVAL_DELAY)) + geom_bar( stat = "identity", alpha = 0.4)

```{r ------------------ Distance }

dis = t
dis = dis %>% mutate(DISTANCE = cut(DISTANCE, breaks = seq(1, max(DISTANCE) + 10, 300))) %>% group_by(DISTANCE,ARRIVAL_DELAY) %>% summarise(freq = n())

frq_dis = rep(0,length((dis$DISTANCE)))


for (i in unique(dis$DISTANCE)) {
  if (is.na(i)) next
  s = sum(dis[dis$DISTANCE == i, ]$freq, na.rm = T)
  
  s_0 = dis[dis$DISTANCE == i & dis$ARRIVAL_DELAY == 0, ]$freq[1] / s
  s_1 = dis[dis$DISTANCE == i & dis$ARRIVAL_DELAY == 1, ]$freq[1] / s
  
  if (!is.na(s_0))
    frq_dis[dis$DISTANCE == i & dis$ARRIVAL_DELAY == 0] = s_0
  
  if (!is.na(s_1))
    frq_dis[dis$DISTANCE == i & dis$ARRIVAL_DELAY == 1] = s_1
  
}

dis['frq_dis'] = frq_dis

g3= ggplot(dis, aes(x = DISTANCE, y = frq_dis, fill = ARRIVAL_DELAY)) + geom_bar( stat = "identity", alpha = 0.4)+
  geom_text(aes( label = format(round(frq_dis, 2), nsmall = 2)), vjust = -0.2))
g3
```


---
title: "flightDelay.R"
author: "Joshua"
date: "Wed Apr 26 06:49:39 2017"
---
